

%%% Choose a language %%%

\newif\ifEN
\ENtrue   % uncomment this for english
%\ENfalse   % uncomment this for czech

%%% Configuration of the title page %%%

\def\ThesisTitleStyle{mff} % MFF style
%\def\ThesisTitleStyle{cuni} % uncomment for old-style with cuni.cz logo
%\def\ThesisTitleStyle{natur} % uncomment for nature faculty logo

\def\UKFaculty{Faculty of Mathematics and Physics}
%\def\UKFaculty{Faculty of Science}

\def\UKName{Charles University in Prague} % this is not used in the "mff" style

% Thesis type names, as used in several places in the title
%\def\ThesisTypeTitle{\ifEN BACHELOR THESIS \else BAKALÁŘSKÁ PRÁCE \fi}
\def\ThesisTypeTitle{\ifEN MASTER THESIS \else DIPLOMOVÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN RIGOROUS THESIS \else RIGORÓZNÍ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN DOCTORAL THESIS \else DISERTAČNÍ PRÁCE \fi}
%\def\ThesisGenitive{\ifEN bachelor \else bakalářské \fi}
\def\ThesisGenitive{\ifEN master \else diplomové \fi}
%\def\ThesisGenitive{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisGenitive{\ifEN doctoral \else disertační \fi}
%\def\ThesisAccusative{\ifEN bachelor \else bakalářskou \fi}
\def\ThesisAccusative{\ifEN master \else diplomovou \fi}
%\def\ThesisAccusative{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisAccusative{\ifEN doctoral \else disertační \fi}



%%% Fill in your details %%%

% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{Performance of Open vSwitch-based Kubernetes Cluster in Pathological Cases}
\def\ThesisAuthor{Václav Šraier}
\def\YearSubmitted{2023}

% department assigned to the thesis
\def\Department{Department of Distributed and Dependable Systems}
% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

\def\Supervisor{Jiří Benc}
\def\SupervisorsDepartment{Red Hat Czech s.r.o.}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Software Systems}

\def\Dedication{%
Dedication. \xxx{It is nice to say thanks to supervisors, friends, family, book authors and food providers.}
}

\def\AbstractEN{%
With the adoption of cloud computing, horizontally scalable infrastructure, and containerized deployments, Software Defined Networking (SDN) became an integral part of data centers, Kubernetes and Open vSwitch (OVS) being one of the commonly deployed solutions. Our work explores the possible performance limitations of OVS under Kubernetes, focusing on pathological traffic patterns. We discovered several types of packets causing excess system load on the cluster nodes. We identified the root cause as a series of drop rules in OpenFlow and a design flaw in OVS that prevents their efficient evaluation. We investigated the impact of this problem and our research revealed a specific system configurations under which an adversary can use the discovered inefficiencies for a practical denial of service attack on the local cluster node, bringing the whole networking stack down for all neighbouring containers.
% ABSTRACT IS NOT A COPY OF YOUR THESIS ASSIGNMENT!
}

\def\AbstractCS{%
S nástupem cloud computingu, kontejnerů a horizontálně škálovatelné infrastruktury, se nedílnou součástí datových center staly softwarově definované sítě (SDN). Jedním z běžně nasazovaných řešení je Kubernetes a Open vSwitch (OVS). V této diplomové práci hledáme možná výkonnostní omezení OVS při použití v rámci Kubernetes. Zaměřujeme se na problémy způsobené neobvyklým síťovým provozem. Výsledkem je objev několika typů paketů způsobujících nadměrné zatížení uzlů clusteru. Jako hlavní příčinu jsme identifikovali řadu filtračních pravidel v OpenFlow a chybu v návrhu OVS, která brání jejich efektivnímu vyhodnocování. Při specifické konfiguraci systému toto potenciálním útočníkům umožňuje využít objevenou neefektivitu k praktickému Denial-of-Service útoku na místní uzel clusteru, který způsobí kompletní síťový výpadek pro všechny kontejnery.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces.
% Keywords are useful for indexing and searching for the theses by topic.
\def\Keywords{%
{Kubernetes} {Open Virtual Network (OVN)} {Open vSwitch (OVS)} {OVN-Kubernetes} {Linux} {networking}
}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
% Alternatively, consider increasing the size of the page by uncommenting the
% geometry modification in thesis.tex.
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  %uncomment to decrease font size

\ifEN\relax\else
% If you are writing a czech thesis, you additionally need to fill in the
% english translation of the metadata here!
\def\ThesisTitleEN{\xxx{Thesis title in English}}
\def\DepartmentEN{\xxx{Name of the department in English}}
\def\DeptTypeEN{\xxx{Department}}
\def\SupervisorsDepartmentEN{\xxx{Superdepartment}}
\def\StudyProgrammeEN{\xxx{study programme}}
\def\StudyBranchEN{\xxx{study branch}}
\def\KeywordsEN{%
\xxx{{key} {words}}
}
\fi
